#440 Reproducibility and Statistics Assignment Homework

<!-- This is an Rmd comment. -->
<!--
You can make comments several lines long by inserting
the text in between the the start and the end of the 
comment.

Comments are where I'll tell you what to do. You should
use the space directly below the comments to do what
the comment asks. You should also describe what you're
doing to make it obvious to the reader so they can
easily follow your logic in a fully reproducible way.
For any statistical tests you run, you should explain
the results in plain English and present the statistics
as you would in an APA style manuscript. For any symbols
you don't know how to use/can't use, feel free to use the name
instead (e.g. alpha)
--> 

<!-- Load the politics.csv data file. -->
First we need to load the data.

```{r}
pols<-read.csv("politics.csv")
```

<!-- 
Check the structure.
subject, party, testtime, minwage, and sex need to be factors.
optimismscore and income should be integer and numeric vectors.
Fix any variables that aren't right.
-->
Let's take a look at the data.
```{r}
str(pols)
```
Now we need to make sure subject, party, testtime, minwage, and sex are factors. Optimismscore and income should be integer and numeric vectors. So let's fix those real quick.
```{r}
pols$subject<-factor(pols$subject)
pols$testtime<-factor(pols$testtime, levels=c("pre", "post"))
```
Okay lets double check the information to make sure our factors changed
```{r}
str(pols)
```

<!-- Find the minimum, mean, and variance for posttest
optimism scores. -->
Next, I'd like to find the minimum, mean, and median incomes. 
```{r}summary(politics$optimismscore[politics$testtime=="post"])
```
We will also find the variance.
```{r}
var(pols$optimism[(pols$testtime="post")])
```

<!-- Create a histogram of posttest optimism scores.
Feel free to use the default ugly settings since
we didn't cover how to do this in ggplot2. If
you're a real go-getter, see if you can figure
out how to make a prettier histogram with better
labels, etc. using ggplot2. -->
Let's see how posttest optimism scores are distributed. To do this we will creat a histogram.
```{r}
hist(pols$optimismscore[pols$testtime=="post"])
```

<!-- Run a test to determine whether political affiliation
is independent of one's gender.
Present the data supporting your analyses in graphical or
tabular form.
-->
Let's see if political affiliation is independent of one's gender. To do this I need to calculate the frequency that individuals appear in different political groups. Then I'll
use a Chi-Square test of independence to test the hypothesis that
affiliation and gender are independent. Again, I don't want
to count people twice, so I'll focus on the posttest data.
```{r}
tab<-table(pols$party,pols$sex)
tab
```
Great, now lets run the Chi-Square test.
```{r}
chisq.test(pols$party,pols$sex)
```
Polical affiliation is not independent of one's gender,Chi-Square(2)=1.45, p=.48

<!-- Run a test to determine whether or not males and females
have different incomes.
Create a graph to display your data that includes the relevant
means and standard errors.-->
Now let's see whether or not males and females have different incomes. To do this we can run an independent t-test.
```{r}
t.test(pols$income[pols$sex=="male"],pols$income[pols$sex=="female"],paired=FALSE,var.equal=FALSE,alternative="two.sided",con.level=.95)
```

To create the figure, I need to include the dplyr library. Then
I'll use it to summarize the politics data set and find the means
and standard errors for the different conditions. Then I'll need
the ggplot2 library to create a nice plot that's easy to work with.
I'll do the same thing for any other graphs I create unless I
specify otherwise.
```{r}
library("dplyr")
temp<-pols[pols$testtime=="post",]%>%group_by(sex)%>%summarize(means=mean(income),
              sems=sd(income)/sqrt(length(income)))
library("ggplot2")
f<-ggplot(temp,aes(x=factor(sex),y=means))+
    geom_bar(stat="identity", color="black",fill=c("deeppink","deepskyblue2"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems), width=.1)
f<-f+ggtitle("Income by Sex")+
    labs(x="Sex", y="Income\n(higher=more income")+
    scale_x_discrete(breaks=c("female","male"),labels=c("Female","Male"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(), axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())
f
```

<!-- See if male and females with different political affiliations
have different posttest optimism scores. Create a graph to display
your data that includes the relevant means and standard errors.-->
Let's see if party affiliation and sex predict posttest optimism scores independently
or in an interaction using a 2-way between-subjects ANOVA.
```{r}
summary(aov(optimismscore~party*sex,data=pols[pols$testtime=="post",]))
```


To create the plot and manually adjust the fill, I need to use
hexadecimal color codes. To convert the named colors I like
to hexadecimal, I need to use the gplots library.
```{r}
temp<-pols[pols$testtime=="post",]%>%group_by(party,sex)%>%
    summarize(means=mean(optimismscore),sems=sd(optimismscore)/sqrt(length(optimismscore)))
library("gplots")
col1=col2hex("deeppink")
col2=col2hex("deepskyblue2")
f<-ggplot(temp, aes(x=party, y=means, fill=sex))+
    geom_bar(stat="identity",position=position_dodge())+
    scale_fill_manual(values=c(col1,col2),name="Sex",breaks=c("female","male"),labels=c("Female", "Male"))+
    theme(legend.key=element_rect(color="black"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2,position=position_dodge(.9))+
    ggtitle("Posttest Optimism score by Sex and Political Affiliation")+
    labs(x="Political Party Affiliation",y="Posttest Optimism Score\n(tens)")+
    scale_x_discrete(breaks=c("democrat","independent","republican"),labels=c("Democrat","Independent","Republican"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(),axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())+
    theme(legend.position=c(.9,.80))+
    theme(legend.background=element_blank())+
    theme(legend.background=element_rect(color="black"))+
    theme(legend.title=element_blank())+
    theme(legend.title=element_text(size=12))+
    theme(legend.title.align=.5)+
    theme(legend.text=element_text(size=10,face="bold"))
f
```



<!-- Create a line graph for the same data. With
error bars of course!-->
Now we will create a line graph for extra practice.
```{r}
f<-ggplot(temp, aes(x=party, y=means, group=sex, color=sex))+
    geom_line(size=1)+
    geom_point(size=2)+
    scale_color_manual(values=c(col1,col2),name="Sex",breaks=c("female","male"),labels=c("Female", "Male"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2)+
    ggtitle("Posttest Optimism Score by Sex and Political Affiliation")+
    labs(x="Political Party Affiliation",y="Posttest Optimism Scores\n(tens)")+
    scale_x_discrete(breaks=c("democrat","independent","republican"),labels=c("Democrat","Independent","Republican"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(),axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())+
    theme(legend.position=c(.9,.80))+
    theme(legend.background=element_blank())+
    theme(legend.background=element_rect(color="black"))+
    theme(legend.title=element_blank())+
    theme(legend.title=element_text(size=12))+
    theme(legend.title.align=.5)+
    theme(legend.text=element_text(size=10,face="bold"))
f
```

<!-- Run a test to see if optimism scores pre- and post
watching videos depends on sex.-->
To see if optimism scores before and after watching *AFHV*
vary depending on gender, I'm going to run a
2-way mixed ANOVA. I'm using a mixed ANOVA because the same
people indicated their optimism before and after the videos
(i.e. a within-subjects factor) and because different
people are different genders (i.e.
a between-subjects factor).
(P.S. -Professor Knapp please see Email for this problem)
```{r}

```

<!-- Run a Multiple Regression Analysis to see
whether you can predict someones posttest optimism
scores from their pretest optimism scores and party
affiliation. Create a figure that shows the distributions
of pre- and posttest optimism scores and sex and create
lines that best fit the data for both sexes regardless of
whether the lines are warranted based on the statistical
analyses-->
I'm going to perform Multiple Regression
to predict prosttest optimism scores from pretest optimism scores and
party affiliation.
```{r}

```
***Fin***